---
title: "Data Science 2401 - Final Project"
author: "Placido Garay"
date: "2023-11-27"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
install.packages("kableExtra", repos = "http://cran.us.r-project.org")
#used this to tidy up the data
library(tidyverse)
#used for data manipulation
library(dplyr)
#used for data visualization
library(ggplot2)
#Scroll boxes
library(kableExtra)
```

## Introduction

[Sport Dataset Link](https://data.world/coreyhermanson/toughest-sport-by-skill)

This data set conducts 60 sports that basis their athleticism in 10 categories. The 10 categories is then totaled and ranked based on how difficult the sport is by skill. The data was calculated by a group of 8 sports scientists from the United States Olympic Committee. There isn't a definitive timeline of when the data was calculated, but was published in 2017.


```{r, include=FALSE}
library(readxl)
toughestsport <- read_excel("C:/Users/Placido Garay/OneDrive - University of Houston Downtown/Desktop/R files/Final Project/toughestsport.xlsx")
View(toughestsport)

```

## Questions to ask
- What and how does the unfiltered data-set explain itself?
- How does each sport rank respically within their own sport category that include:

- 1. Toughest Snow/Ice Sports
- 2. Toughest Adventure Sports
- 3. Toughest Water Sports
- 4. Toughest Team Sport
- 5. Toughest Individual Sport

- For each data-set, what is the max,min and mean of difficulty?
- What are the easiest sports ranked by difficulty?

## Packages Used
```{r, echo = T}
#used this to tidy up the data
library(tidyverse)
#used for data manipulation
library(dplyr)
#used for data visualization
library(ggplot2)
#Scroll boxes
library(kableExtra)
```

## Unfiltered Dataset

```{r, echo=TRUE}
glimpse(toughestsport)
```

## Key Terms
- END (Endurance): The ability to continue to perform a skill or action for long periods of time.
- STR (Strength): The ability to produce force.
- PWR (Power): The ability to produce strength in the shortest possible time.
- SPD (Speed): The ability to move quickly.
- AGI(Agility): The ability to change direction quickly.
- FLX (Flexibility): The ability to stretch the joints across a large range of motion.
- NER (Nerve): The ability to overcome fear.
- DUR(Durability): The ability to withstand physical punishment over a long period of time.
- HAN (Hand-Eye Coordination): The ability to react quickly to sensory perception.
- ANA (Analytic Aptitude): The ability to evaluate and react appropriately to strategic situations.



## Calculations
- Total # of sports: 60
- Categories (Skills): 10
- Out of the 10 categories, skills are arranged from a 1-10 scale based on the demands for each sport.
- Total # of skills: All 10 categories are added and are based individually for each sport.
- Rank: All 60 sports are ranked based on their degree of difficulty total amount.




## Unfiltered Plot/Data
```{r, echo=FALSE}
toughestsport %>% 
  ggplot()+aes(x=reorder(SPORT,-RANK), y=TOTAL, fill=TOTAL)+geom_col()+ labs(title = 'Degree of difficulty in 60 sports',x='Sports',y='Total')+coord_flip()+ geom_text(aes(label=TOTAL,size=2.5))+scale_size_identity()


toughestsport %>% kbl() %>% kable_paper()  %>% scroll_box(height = "800px")

toughestsport %>% arrange(SPORT) %>% 
  summarise(max(TOTAL),min(TOTAL),mean(TOTAL))
toughestsport
```


## Toughest Snow/Ice Sports
```{r, echo=FALSE}
ice_sports <- toughestsport %>% 
  filter(SPORT=='Ice Hockey'| SPORT=='Skiing: Alpine'| SPORT=='Speed Skating'| SPORT=='Figure Skating'| SPORT=='Skiing: Freestyle'| SPORT=='Bobsledding/Luge'| SPORT=='Ski Jumping'| SPORT=='Skiing: Nordic'| SPORT=='Curling') %>% 
  select(SPORT,TOTAL) %>%  mutate(RANK2=dense_rank(desc(TOTAL)))


ice_sports %>% 
   ggplot()+aes(x=reorder(SPORT,-RANK2), y=TOTAL, fill=TOTAL)+geom_col()+ labs(title = 'Degree of difficulty Snow/Ice Sports',x='Sports',y='Total')+coord_flip()+geom_text(aes(label=TOTAL),position = position_jitter(width = 0))


ice_sports %>% arrange(SPORT) %>% 
  summarise(max(TOTAL),min(TOTAL),mean(TOTAL)) 
ice_sports
 

```


## Toughest Adventure Sports
```{r, echo=FALSE}
adv_sports <- toughestsport %>% 
  filter(SPORT=='Surfing'| SPORT=='Skateboarding'| SPORT=='Rowing'| SPORT=='Water Skiing'| SPORT=='Canoe/Kayak'| SPORT=='Horse Racing'| SPORT=='Cheerleading'| SPORT=='Equestrian'| SPORT=='Shooting'| SPORT=='Fishing') %>% 
  select(SPORT,TOTAL)%>%  mutate(RANK3=dense_rank(desc(TOTAL)))


adv_sports %>% 
   ggplot()+aes(x=reorder(SPORT,-RANK3), y=TOTAL, fill=TOTAL)+geom_col()+ labs(title = 'Degree of difficulty Outdoor-Adventure Sports',x='Sports',y='Total')+coord_flip()+geom_text(aes(label=TOTAL),position = position_jitter(width = 0))

adv_sports %>% arrange(SPORT) %>% 
  summarise(max(TOTAL),min(TOTAL),mean(TOTAL))
adv_sports
```

## Toughest Water Sports
```{r, echo=FALSE}
water_sports <- toughestsport %>% 
  filter(SPORT=='Water Polo'| SPORT=='Diving'| SPORT=='Swimming (all strokes): Distance'| SPORT=='Water Skiing'| SPORT=='Swimming (all strokes): Sprints'| SPORT=='Fishing') %>% 
  select(SPORT,TOTAL)%>%  mutate(RANK4=dense_rank(desc(TOTAL)))


water_sports %>% 
   ggplot()+aes(x=reorder(SPORT,-RANK4), y=TOTAL, fill=TOTAL)+geom_col()+ labs(title = 'Degree of difficulty Water Sports',x='Sports',y='Total')+coord_flip()+geom_text(aes(label=TOTAL),position = position_jitter(width = 0))

water_sports %>% arrange(SPORT) %>% 
  summarise(max(TOTAL),min(TOTAL),mean(TOTAL))
water_sports
```


## Toughest Team Sport
```{r, echo=FALSE}
team_sports <- toughestsport %>% 
  filter(SPORT=='Ice Hockey'| SPORT=='Football'| SPORT=='Basketball'| SPORT=='Baseball/Softball'| SPORT=='Soccer'| SPORT=='Water Polo'| SPORT=='Rugby'| SPORT=='Lacrosse'| SPORT=='Volleyball'| SPORT=='Team Handball'| SPORT=='Cheerleading') %>% 
  select(SPORT,TOTAL)%>%  mutate(RANK4=dense_rank(desc(TOTAL)))


team_sports %>% 
   ggplot()+aes(x=reorder(SPORT,-RANK4), y=TOTAL, fill=TOTAL)+geom_col()+ labs(title = 'Degree of difficulty Team Sports',x='Sports',y='Total')+coord_flip()+geom_text(aes(label=TOTAL),position = position_jitter(width = 0))

team_sports %>% arrange(SPORT) %>% 
  summarise(max(TOTAL),min(TOTAL),mean(TOTAL))
team_sports

```

## Toughest Individual Sport
```{r, echo=FALSE}
Indv_sports <- toughestsport %>% 
  filter(SPORT=='Wrestling'| SPORT=='Martial Arts'| SPORT=='Tennis'| SPORT=='Gymnastics'| SPORT=='Skiing: Alpine'| SPORT=='Rodeo: Steer Wrestling'| SPORT=='Speed Skating'| SPORT=='Figure Skating'| SPORT=='Volleyball'| SPORT=='Cycling: Distance'| SPORT=='Racquetball/Squash'| SPORT=='Track and Field: Pole Vault'| SPORT=='Surfing'| SPORT=='Fencing'| SPORT=='Cycling: Sprints'| SPORT=='Ski Jumping'| SPORT=='Badminton'| SPORT=='Auto Racing'| SPORT=='Diving'| SPORT=='Weight-Lifting'| SPORT=='Table Tennis'| SPORT=='Horse Racing'| SPORT=='Golf'| SPORT=='Archery'| SPORT=='Bowling'| SPORT=='Shooting'| SPORT=='Fishing'| SPORT=='Boxing') %>% 
  select(SPORT,TOTAL) %>%  mutate(RANK5=dense_rank(desc(TOTAL)))


Indv_sports %>% 
   ggplot()+aes(x=reorder(SPORT,-RANK5), y=TOTAL, fill=TOTAL)+geom_col()+ labs(title = 'Degree of difficulty Individual Sports',x='Sports',y='Total')+coord_flip()+geom_text(aes(label=TOTAL),position = position_jitter(width = 0))

 Indv_sports %>% kbl() %>% kable_paper()  %>% scroll_box(height = "800px")
Indv_sports %>% arrange(SPORT) %>% 
  
  summarise(max(TOTAL),min(TOTAL),mean(TOTAL))
Indv_sports

```
## Easiest Sports Ranked
```{r, echo=FALSE}

easiest_sport<- toughestsport %>% select(SPORT,END,STR,PWR,SPD,AGI,FLX,NER,DUR,HAN,ANA,TOTAL) %>%  mutate(NEW_RANK=dense_rank(TOTAL)) %>% arrange(NEW_RANK)

easiest_sport %>% 
  ggplot()+aes(x=reorder(SPORT,-NEW_RANK), y=TOTAL, fill=TOTAL)+geom_col()+ labs(title = 'Degree of simplicity in 54 sports',x='Sports',y='Total')+coord_flip()+ geom_text(aes(label=TOTAL,size=2.5))+scale_size_identity()


easiest_sport %>% 
 kbl() %>% kable_paper()  %>% scroll_box(height = "800px")

easiest_sport %>% arrange(SPORT) %>% 
  summarise(max(TOTAL),min(TOTAL),mean(TOTAL))
```
## Conclusion

- In conclusion, we can determine that the data-set reveals that out of 60 sports. Boxing is considered the most difficult sport and fishing is more effortless. Team sports seemed to be more competitive and difficult compared to individual sports, based on their mean total. 


